% =====================================================================
%  PAPER 3  —  A curvature diagnostic for continuous-RSB criticality
%             versus kinematic volume-divergence in disordered systems
%  IG-PRIMON-T1.  Draft v0.1 (2026-06-14).
%  Compilable: pdflatex twice (refs/labels).
%  STATUS (HONEST_CLAIMS):
%    - ridge kinematic side .......... [V]  (module_L_ridge_curvature.py)
%    - SK genuine side (R_eps) ....... [V]  (module_L_SK_converse.py)
%    - perceptron susceptibility ..... [V]  (module_L_perceptron_replicon.py)
%    - perceptron CURVATURE .......... [V]  (module_L_perceptron_finiteT.py, positive-definite
%                                            (beta,eps); indefinite (alpha,eps) first pass =
%                                            module_L_perceptron_curvature.py, recorded as such)
%  Citations verified against sources 2026-06-14 (Gardner, Gardner-Derrida, Krauth-Mezard,
%  Gyorgyi, de Almeida-Thouless, Sherrington-Kirkpatrick, Ruppeiner, Janyszek-Mrugala, Belkin).
%  Related-work positioning + Ersoy-Wiesner (arXiv:2505.06597) and Ziyin-Ueda (PRR 5,043243,2023)
%  added and verified against the sources 2026-06-15 (EW retitled v1->v2; the second/first-order-
%  by-depth result is Ziyin-Ueda's, which EW reproduce and extend).
%  For J. Stat. Mech. submission, swap \documentclass to {iopart}.
% =====================================================================
\documentclass[a4paper,11pt]{article}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs}
\usepackage[margin=1in]{geometry}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\newcommand{\Ric}{R}
\newcommand{\eps}{\varepsilon}
\newcommand{\detg}{\det g}
\newcommand{\repl}{\lambda_{\mathrm{repl}}}
\DeclareMathOperator{\Dz}{D}

\title{A curvature diagnostic for continuous-RSB criticality versus\\
       kinematic volume-divergence in disordered systems}
\author{Aaron Jones\\ \small JtechAI LLC}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
A diverging response is ambiguous: it can mark genuine interacting criticality, or a purely \emph{kinematic}
explosion of a metric \emph{volume} while nothing critical happens. We give a sharp information-geometric
criterion that separates them, built on the Ruppeiner scalar curvature $\Ric$: a genuine transition is a
diverging metric \emph{component} on a non-degenerate, \emph{positive-definite} metric, whereas a
$\detg\to0$ or coordinate-degenerate divergence is spurious and a manifold lacking the order parameter's
conjugate field is blind. ``$|\Ric|\to\infty$'' alone proves nothing. We validate the criterion on three
exactly-solvable archetypes from distinct venues. (i)~\emph{Learning}: ridge regression at the interpolation
threshold---double descent is kinematic, $\detg\sim\lambda^{-3/2}$ while $\Ric$ stays flat-to-bounded.
(ii)~\emph{Spin glass}: Sherrington--Kirkpatrick approached along the de Almeida--Thouless line, where
$|\Ric|\sim\chi_{\mathrm{SG}}^2$ diverges on a positive-definite metric. (iii)~\emph{Storage/jamming}: the
spherical perceptron, which carries \emph{both} sides in one model---convex storage is kinematic, while
non-convex (negative-margin) storage is a continuous replicon at which $|\Ric|\sim\chi^2$ diverges on a
positive-definite finite-temperature metric, a signature the natural (structural) coordinate fails to
reveal. The phases are classical; the contribution is their uniform curvature \emph{classification} across
venues---learning, spin glasses, and constraint satisfaction---of one subject: continuous-RSB criticality
versus kinematic volume-divergence in disordered systems. We further establish a topological trichotomy for
the genuine signature and argue that for connected, over-parametrized network posteriors it is
vacuous---routing such architectures from the curvature diagnostic to the local learning coefficient.
\end{abstract}

% =====================================================================
\section{Introduction}\label{sec:intro}
The Ruppeiner scalar curvature $\Ric$ of a thermodynamic state space is a geometric measure of fluctuations,
diverging at ordinary critical points where the correlation volume becomes macroscopic
\cite{ruppeiner1995,janyszek1990}. One-dimensional curvature claims are vacuous; on a two-parameter
manifold $\Ric$ is a non-trivial invariant.

A recurring confusion motivates this paper. Modern high-dimensional inference exhibits divergences---most
famously the double-descent peak of the generalization error at the interpolation threshold
\cite{belkin2019}---that \emph{look} like phase transitions but need not be. The question ``is this
divergence a genuine transition?'' has, we argue, a clean geometric answer, and the answer is not ``does
something blow up.'' A metric \emph{volume} $\detg$ can diverge for purely kinematic reasons while the
intrinsic curvature stays bounded; conversely, $|\Ric|\to\infty$ on a degenerate metric can be a coordinate
artifact. The diagnostic must therefore be \emph{refined}.

We make this precise (Sec.~\ref{sec:diagnostic}) and validate it on three exactly-solvable archetypes
(Secs.~\ref{sec:ridge}--\ref{sec:perceptron}). The archetypes are deliberately drawn from three different
\emph{venues}---supervised learning, mean-field spin glasses, and constraint satisfaction/jamming---because
our central claim is that the dichotomy is not about any one of them. ``The geometry of learning
transitions'' would name the venue, not the subject. The subject is the dichotomy between
\emph{continuous-RSB criticality} and \emph{kinematic volume-divergence in disordered systems}.

\paragraph{Status of the claims.} We follow a strict claim hygiene. All three archetypes are verified
($[V]$). For the perceptron (Sec.~\ref{sec:perceptron}) we keep separate two statements that are easily
conflated: the divergence of the \emph{overlap susceptibility} (Sec.~\ref{sec:perc-chi}) and the literal
\emph{curvature} statement $|\Ric|\to\infty$ (Sec.~\ref{sec:perc-R}). The latter is the more demanding
because ``$|\Ric|\sim$ correlation volume'' is Riemannian: it is established here on the
finite-temperature $(\beta,\eps)$ manifold, where both coordinates are conjugate fields and the metric is
positive-definite by construction. A first computation tuned by the structural load $\alpha$ gives the same
divergence on an \emph{indefinite} metric and is recorded as the cautionary intermediate it is.

% =====================================================================
\section{The refined diagnostic}\label{sec:diagnostic}
Let $\psi(\theta^1,\theta^2)=\log Z$ be the log-partition function of a two-parameter exponential family in
natural parameters $\theta=(\theta^1,\theta^2)$, with sufficient statistics $T=(T_1,T_2)$. The Fisher
information / thermodynamic metric is the Hessian
\begin{equation}
g_{ij}=\partial_{\theta^i}\partial_{\theta^j}\psi=\mathrm{Cov}(T_i,T_j),
\label{eq:metric}
\end{equation}
positive-semidefinite by construction. The Ruppeiner scalar curvature of \eqref{eq:metric} is, in the
program's sphere-positive convention,
\begin{equation}
\Ric=-\frac{N}{2(\detg)^2},
\label{eq:R}
\end{equation}
with $N$ the standard determinant of second and third derivatives of $\psi$ (Appendix; the engine is pinned
to $\Ric=-1$ on the normal family). The sign of $\Ric$ is convention-dependent; only the \emph{invariant}
distinction between $|\Ric|$ bounded and $|\Ric|\to\infty$ carries physical content. Equation~\eqref{eq:R}
already exposes the logical engine of the diagnostic: a \emph{diverging} $\detg$ \emph{suppresses} $|\Ric|$
(the kinematic mode---a volume that explodes faster than the numerator), whereas a \emph{vanishing} $\detg$
\emph{inflates} it (the spurious mode); only a diverging numerator component on a bounded-below, positive
$\detg$ produces a genuine divergence.

\begin{definition}[Refined criticality diagnostic]\label{def:diag}
A divergence is a \emph{genuine continuous transition} iff $|\Ric|\to\infty$ is produced by a diverging
metric \emph{component} $g_{\eps\eps}\to\infty$ on a metric that remains \emph{non-degenerate and
positive-definite} ($\detg>0$, well-conditioned), where $\eps$ is the natural field conjugate to the order
parameter. Three configurations fail the criterion --- the first two are disqualified divergences, the
third sees no transition at all:
\begin{itemize}
\item \emph{Kinematic (volume).} $\detg\to\infty$ while $|\Ric|$ stays bounded: the metric volume explodes,
      the intrinsic curvature does not. (Double descent, Sec.~\ref{sec:ridge}.)
\item \emph{Spurious (degenerate).} $|\Ric|\to\infty$ driven by $\detg\to0$ or by functionally dependent
      coordinates (e.g.\ $E\propto M^2$, Curie--Weiss): a coordinate artifact, not a fluctuation.
\item \emph{Blind.} A manifold lacking the order parameter's conjugate field as a coordinate sees no
      transition at all.
\end{itemize}
\end{definition}

The discipline of Definition~\ref{def:diag} is that ``$|\Ric|\to\infty$'' alone proves nothing; the
mechanism and the signature must both be checked.

% =====================================================================
\section{Kinematic archetype: ridge regression and double descent}\label{sec:ridge}
Consider the Gibbs measure over student weights of a linear (Gaussian) teacher--student problem,
$P(w)\propto\exp\!\big(-\beta E-\beta\lambda\,\tfrac12\|w\|^2\big)$ with $E=\tfrac12\|y-Xw/\sqrt N\|^2$ and a
(possibly noisy) teacher $y=Xw_\star/\sqrt N+\sigma\xi$. The sufficient statistics are $(E,\tfrac12\|w\|^2)$
and the natural parameters $(\theta^1,\theta^2)=(-\beta,-\beta\lambda)$---the faithful analog of a gas's
$(-\beta,\log z)$, with the ridge field $\lambda$ in the fugacity role.

\subsection{The exact manifold}\label{sec:ridge-manifold}
Because $E$ is quadratic in $w$, $\log Z$ is a Gaussian integral, diagonal in the eigenbasis $\{v_i\}$ of the
sample covariance $M=X^\top X/N$ with eigenvalues $\{\mu_i\}$ and teacher projections
$d_i=\langle v_i,\,X^\top y/\sqrt N\rangle$:
\begin{equation}
\log Z(\theta)=\mathrm{const}-\tfrac12\textstyle\sum_i\log k_i+\tfrac{\beta^2}{2}\sum_i\frac{d_i^2}{k_i},
\qquad k_i=\beta(\mu_i+\lambda).
\label{eq:ridgeZ}
\end{equation}
The metric \eqref{eq:metric} and all third derivatives are therefore closed-form spectral sums; the
fluctuation that controls the transition is the variance term $g\supset\sum_i k_i^{-2}$. The interpolation
threshold $\alpha=P/N=1$ is the candidate transition: there the Marchenko--Pastur (MP) soft edge touches
zero, so the spectral density behaves as $\rho(\mu)\sim(\pi\sqrt\mu)^{-1}$ as $\mu\to0$, and as the ridge
$\lambda\to0$ the softest mode $k_{\min}=\beta(\mu_{\min}+\lambda)\to0$.

\subsection{The kinematic verdict}\label{sec:ridge-verdict}
At $\beta=1,\alpha=1$ the limit $\lambda\to0$ produces a textbook-looking divergence that the diagnostic
classifies as \emph{kinematic} (receipt \texttt{module\_L\_ridge\_curvature.py}; analytic Hessian validated
against finite differences of $\log Z$ to max.\ relative error $8.9\times10^{-6}$, engine pinned to
$\Ric=-1$ on the normal family):

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
quantity & as $\lambda\to0$ & reading\\
\midrule
metric volume $\detg$ & diverges $\sim\lambda^{-3/2}$ & MP soft-edge fluctuation explosion\\
metric rank & collapses to rank 1 (defect $\to1.000$) & one soft mode dominates\\
$\Ric$, noiseless teacher & $\to0$, $\sim\lambda$ ($-2.0\mathrm{e}{-}5\to-8.1\mathrm{e}{-}10$) & asymptotically flat\\
$\Ric$, noisy teacher ($\sigma=0.5$) & bounded $\approx-1\times10^{-4}$, seed-robust & bounded floor, never diverges\\
\bottomrule
\end{tabular}
\end{center}

The diverging volume is a direct soft-edge effect: from \eqref{eq:ridgeZ} the singular metric component is
$\sum_i k_i^{-2}\to N\!\int_0\rho(\mu)\,[\beta(\mu+\lambda)]^{-2}\,d\mu\sim\!\int_0
\mu^{-1/2}(\mu+\lambda)^{-2}\,d\mu\propto\lambda^{-3/2}$, so $\detg\sim\lambda^{-3/2}$ while the metric
collapses onto the single softest mode (the defect $\to1$).

\subsection{Mechanism: a free versus a sourced soft mode}\label{sec:ridge-mech}
The divergence is thus \emph{entirely} a metric-volume effect: $\detg\to\infty$ while $|\Ric|$ is bounded
(noisy) or vanishing (noiseless). It is the exact cancellation of singular powers in $\Ric=-N/(2(\detg)^2)$,
with $N$ and $(\detg)^2$ diverging at \emph{matching} rates, so that $\Ric$ stays finite whenever the
diverging volume is sourced by a free log-determinant sector---the same flat-sector cancellation that keeps
thermodynamic curvature finite at a limiting-temperature transition in product-form gases. The mechanism is
transparent at the soft mode: with a noiseless teacher the soft direction is a \emph{free} log-determinant
mode (signal weight $d^2_{\min}\sim\mu^2_{\min}\to0$ vanishes with the eigenvalue), giving exactly flat
curvature, the classical ideal-gas baseline; noise gives that mode a \emph{source} ($d^2_{\min}\to$ the
noise floor $\sim\sigma^2$), lifting $\Ric$ to a bounded nonzero value. Neither diverges. Forty-digit
arithmetic is required, not optional: the rank-1 collapse costs $\sim11$ digits to cancellation, and at
single precision the bounded $\Ric$ masquerades as a divergence---a standing caution for numerical
thermodynamic geometry in high dimension.

% =====================================================================
\section{Genuine archetype: the SK spin glass at the dAT line}\label{sec:sk}
The Sherrington--Kirkpatrick model \cite{sk1975} is the positive control: a genuine interacting transition,
reached without any replica-symmetry-breaking ansatz.

\subsection{The two-replica construction}\label{sec:sk-construct}
We stay replica-symmetric and \emph{approach} (never cross) the de Almeida--Thouless (dAT) instability
\cite{at1978}. Take two real replicas $s^1,s^2$ and couple them by a field $\eps$ conjugate to their overlap
$O=\sum_i s^1_i s^2_i$; the disorder-averaged two-replica RS free energy $\psi_2(\beta,h,\eps)$ is then a
function on a manifold whose $\eps$-direction \emph{is} the order-parameter channel. At $\eps=0$ the replicas
decouple, $\psi_2=2\psi_1$ (verified to $|\text{diff}|=0$), so $\eps$ is a clean perturbation of the
single-replica state and the $\eps$-metric component is the spin-glass susceptibility
$g_{\eps\eps}=\beta\,\partial p/\partial\eps=\beta\chi_{\mathrm{SG}}$, $p=\langle O\rangle/N$ the
inter-replica overlap.

\subsection{The $h=0$ closed form}\label{sec:sk-h0}
At $h=0$ the construction is fully explicit:
\begin{equation}
\psi_2(\beta,\eps)=\tfrac{\beta^2}{2}(1-p^2)+\log4+\log\cosh(\beta^2 p+\beta\eps),\qquad p=\tanh(\beta^2 p+\beta\eps).
\label{eq:psi2}
\end{equation}
For $\beta<1$ the saddle is $p=0$ at $\eps=0$; differentiating the self-consistency,
$\partial p/\partial\eps|_0=\beta\,\mathrm{sech}^2(0)/(1-\beta^2\mathrm{sech}^2(0))=\beta/(1-\beta^2)$, so
with the replicon $\repl\equiv1-\beta^2$,
\begin{equation}
g_{\eps\eps}=\frac{\beta^2}{\repl},\qquad
\Ric_{\text{bare}}=\frac{4}{\beta^2}\ (\text{bounded: bare manifold \emph{blind}}),\qquad
\Ric_{\eps}=\frac{2}{\beta^2\repl^2}\xrightarrow{\;\repl\to0\;}\infty,\qquad \detg=\frac{\beta^2}{\repl}>0.
\label{eq:sk}
\end{equation}
So $|\Ric_\eps|\sim\chi_{\mathrm{SG}}^2$ diverges on a \emph{strictly positive-definite} metric---a genuine
transition by Definition~\ref{def:diag} (e.g.\ $\Ric_\eps=68,\,233,\,1328$ at $\beta=0.90,0.95,0.98$), the
exact prototype the perceptron reproduces in Sec.~\ref{sec:perc-R}. The bare $(\beta,h)$ manifold, lacking
the $\eps$-coordinate, sees only the bounded $\Ric_{\text{bare}}=4/\beta^2$: it is \emph{blind}.

\subsection{Along the de Almeida--Thouless line}\label{sec:sk-dat}
Off the symmetric axis there is no closed form: the coupled $(p,q)$ saddle is solved on Gauss--Hermite nodes
and $g_{\eps\eps}$ obtained by \emph{implicit differentiation} (linear response). It scales cleanly as
$g_{\eps\eps}=\beta\chi_{\mathrm{SG}}\sim C(h)/\lambda_{AT}$ all along the dAT line, the constant $C(h)$
stabilizing as the replicon $\lambda_{AT}\to0$:

\begin{center}
\begin{tabular}{@{}cccccc@{}}
\toprule
$h$ & $\lambda_{AT}$ & $\beta$ & $g_{\eps\eps}$ (implicit) & $g_{\eps\eps}$ (fin.\ diff.) & $\chi_{\mathrm{SG}}\lambda_{AT}$\\
\midrule
$0.3$ & $0.30$ & $1.013$ & $3.83$ & $3.83$ & $1.13$\\
$0.3$ & $0.10$ & $1.342$ & $17.36$ & $17.98$ & $1.29$\\
$0.3$ & $0.02$ & $1.495$ & $97.46$ & $30.98$ & $1.30$\\
$0.5$ & $0.30$ & $1.265$ & $4.35$ & $4.35$ & $1.03$\\
$0.5$ & $0.10$ & $1.662$ & $17.76$ & $18.29$ & $1.07$\\
$0.5$ & $0.02$ & $1.827$ & $97.79$ & $32.03$ & $1.07$\\
\bottomrule
\end{tabular}
\end{center}

The product $\chi_{\mathrm{SG}}\lambda_{AT}\to C(h)$ flattens to three digits, confirming
$g_{\eps\eps}\sim1/\lambda_{AT}\to\infty$; note the finite-difference column \emph{drifts} below the stable
implicit value as $\lambda_{AT}\to0$ ($30.98$ vs.\ $97.46$ at $\lambda_{AT}=0.02$), the breakdown quantified
in Remark~\ref{rem:fd}. The divergence mechanism is $\det(I-M)\to0$---the two-replica saddle's stability
matrix going singular, which \emph{is} the replicon condition---with $\detg>0$ and growing throughout, while
the bare $(\beta,h)$ curvature stays bounded ($\approx16$ at $h=0.3$): blindness confirmed off-axis. The
diagnostic tracks the replicon wherever it sits in the $(T,h)$ plane
\cite[receipt \texttt{module\_L\_SK\_converse.py}]{repo}.

\begin{remark}[Numerical caveat, honestly recorded]\label{rem:fd}
Third-derivative finite differences of $\Ric$ break down for $\lambda_{AT}\lesssim0.05$ (in double precision
on an \texttt{fsolve}'d saddle, $\Ric$ sign-flips and the finite-difference $g_{\eps\eps}$ itself drifts,
$31$ vs.\ $97$ at $\lambda_{AT}=0.02$). The $|\Ric|\to\infty$ claim therefore rests on the $h=0$ closed
form \eqref{eq:sk}, the stable linear-response $g_{\eps\eps}$ (clean to $\lambda_{AT}=0.02$), and the $\Ric$
growth where the third difference is still reliable ($\lambda_{AT}\gtrsim0.1$: $26\to41\to99$)---never on the
edge values. The same caveat governs the perceptron curvature (Sec.~\ref{sec:perc-R}), which is likewise
built from a soft-mode/linear-response construction rather than naive edge differences.
\end{remark}

% =====================================================================
\section{The third archetype: the spherical perceptron carries both sides}\label{sec:perceptron}
The Gardner storage problem for a perceptron \cite{gardner1988,gardnerderrida1988} is, geometrically, a
single model that realises \emph{both} sides of the dichotomy as the weight geometry is varied. For weights
$w$ ($\|w\|^2=N$ spherical, or $w\in\{\pm1\}^N$ Ising), $p=\alpha N$ random patterns at margin $\kappa$, the
Gardner volume $V=\int d\mu(w)\prod_\mu\Theta(\sigma^\mu(w\!\cdot\!\xi^\mu)/\sqrt N-\kappa)$ has the
replica-symmetric free entropy $s(q)=G_S(q)+\alpha\,G_E(q)$ with the energetic term
\begin{equation}
G_E(q)=\int \Dz t\,\log H(u),\qquad u=\frac{\kappa-\sqrt q\,t}{\sqrt{1-q}},\qquad
H(x)=\tfrac12\operatorname{erfc}(x/\sqrt2),
\label{eq:GE}
\end{equation}
($\Dz t$ the standard Gaussian measure) and the spherical entropic term
$G_S(q)=\tfrac12[\log(1-q)+q/(1-q)]$. We emphasise that \eqref{eq:GE} is the log-volume of a
\emph{constraint-satisfaction} problem; the transition below is jamming, not generalization.

The \emph{phases} of this model are classical: the Gardner capacity \cite{gardner1988,gardnerderrida1988},
the frozen-1RSB capacity of the Ising case \cite{krauthmezard1989}, and the continuous-RSB / jamming
instability of the negative-margin case are all established results. We claim none of them as new. The
contribution of this section is not the phase diagram but its \emph{classification by curvature signature}:
we show that the \emph{same} information-geometric criterion (Definition~\ref{def:diag}) that separates
double descent from SK assigns each of these known phases to the correct tier, and in particular that one
model supplies both a kinematic and a genuine entry. The lens is the object of study; the phases validate
it.

\subsection{Convex storage is kinematic; Ising storage is first-order}\label{sec:perc-convex}
For $\kappa\ge0$ the problem is convex: the RS solution is exact up to capacity and there is no replicon
instability. The only transition is $q\to1$ at the Gardner capacity
$\alpha_c(\kappa)=\big[\int_{-\kappa}^\infty \Dz t\,(t+\kappa)^2\big]^{-1}$ (so $\alpha_c(0)=2$ exactly,
verified), a metric-\emph{volume} effect---the storage twin of double descent, the kinematic side. For
Ising weights the storage transition is instead \emph{frozen-1RSB}: the RS free entropy starts at
$s(0)=\log2$ and goes negative, crossing zero at $\alpha_{RS}=0.83308$, coinciding with the Krauth--M\'ezard
capacity $0.833$ \cite{krauthmezard1989}; a negative quenched entropy for discrete weights is impossible, so
the solution space freezes discontinuously ($q$ jumps) and the continuous-replicon route does not apply.
(The Ising \emph{learning} transition is likewise first order \cite{gyorgyi1990}.)

\subsection{Non-convex storage: susceptibility-genuine $[V]$}\label{sec:perc-chi}
For $\kappa<0$ the problem is non-convex and a continuous replicon appears below capacity. From the RS saddle
$q/(1-q)=\alpha\int\Dz t\,(H'/H)^2$ the replicon (AT) eigenvalue is
\begin{equation}
\repl=1-\alpha\int \Dz t\,\big[G_1'(u)\big]^2,\qquad G_1=H'/H,\quad G_1'=-uG_1-G_1^2.
\label{eq:repl}
\end{equation}
Equation~\eqref{eq:repl} is anchored to Gardner: at $\kappa=0$ it vanishes precisely at $\alpha=2=\alpha_c$
(the AT line meets capacity, marginal). For $\kappa<0$ it crosses zero \emph{continuously} at
$\alpha_{AT}(\kappa)<\alpha_c(\kappa)$ (e.g.\ $\alpha_{AT}=2.32,2.67,3.05,3.95$ for
$\kappa=-0.1,-0.2,-0.3,-0.5$), and the overlap susceptibility $\chi=1/\repl$ diverges continuously,
$\chi\cdot(\alpha_{AT}-\alpha)\to\text{const}$ ($\approx3.22$ at $\kappa=-0.5$)---a clean
$1/(\alpha_{AT}-\alpha)$ power law, the direct twin of SK's $\chi_{\mathrm{SG}}\sim1/\lambda_{AT}$. Since
$\eps$ (conjugate to the overlap) is a genuine natural field, $g_{\eps\eps}\sim1/\repl$ is a true
conjugate-field susceptibility whose divergence is coordinate-independent, Gardner-anchored, and
component-driven. \textbf{By the susceptibility criterion the non-convex spherical perceptron sits on the
genuine side; this is verified} $[V]$ \cite[receipt \texttt{module\_L\_perceptron\_replicon.py}]{repo}.

\subsection{The literal curvature: $[V]$, on a positive-definite manifold}\label{sec:perc-R}
On the RS side the overlap sector is a Gaussian soft mode, so the $\eps$-augmented free energy is
$\psi(c,\eps)=2\,\psi_0(c)+\chi(c)\,\eps^2/2+O(\eps^4)$ in a \emph{generic} tuning coordinate $c$, with
$\chi=1/\repl$ the overlap susceptibility. We instantiate $c$ two ways below---first the structural load
$\alpha$, which fails the signature test, then the natural field $\beta$, which passes---to show the
diagnostic is doing real work.

At $\eps=0$ the engine inputs are $g_{\eps\eps}=\chi$ (singular), $g_{cc}=2\psi_0''$ and $g_{c\eps}=0$
(regular), and among the third derivatives only the cross-term $\partial_c g_{\eps\eps}=\chi'\sim\chi^2$ is
singular (since $c$ tunes $\repl$), while $\partial_\eps^3\psi=0$ and the rest are $O(1)$. The numerator $N$
of \eqref{eq:R} is then dominated by the single term $\propto(\partial_c g_{\eps\eps})^2\sim\chi^4$, whereas
$\detg\sim g_{cc}\,\chi\sim\chi$; hence $\Ric=-N/(2(\detg)^2)\sim\chi^4/\chi^2=\chi^2$, i.e.
$|\Ric|\sim\chi^2\sim1/(c_{AT}-c)^2$, the curvature diverging as the \emph{square} of the susceptibility,
exactly as SK's $\Ric_\eps\sim\chi_{\mathrm{SG}}^2$. Crucially this is a genuine divergence, \emph{not} the
ridge cancellation of Sec.~\ref{sec:ridge}: there the singular powers of $N$ and $(\detg)^2$ matched and
cancelled, whereas here the singular cross-derivative $\partial_c g_{\eps\eps}$ survives uncancelled.
Freezing $\chi$ (i.e.\ setting $\chi'=0$) sends $\Ric\to0$, confirming the divergence is a \emph{component}
effect, not a background one.

The signature, however, depends on \emph{which} tuning coordinate is used, and this is where
Definition~\ref{def:diag} earns its keep. Tuning with the load $\alpha$ (a \emph{structural} parameter, not a
natural field) gives $g_{\alpha\alpha}=2s''<0$ (the free entropy is concave): $|\Ric|\to\infty$ but on an
\emph{indefinite} metric ($\detg<0$), which does not establish the Riemannian statement
``$|\Ric|\sim$ correlation volume.'' We therefore tune instead with a genuine natural field. At finite
temperature the storage problem is a Gibbs measure with inverse temperature $\beta$ conjugate to the
constraint energy; the soft constraint $a=e^{-\beta}$ gives $m=a+(1-a)H(u)$ and $M_1=(1-a)H'/m$, with the
replicon retaining its form, $\repl=1-\alpha\int\Dz t\,[M_1']^2$, $M_1'=-uM_1-M_1^2$ (recovering
\eqref{eq:repl} as $\beta\to\infty$). Now $\beta$ is a natural field, so $g_{\beta\beta}=2\psi_0''=
2\,\mathrm{Var}(\text{energy})/N>0$, \emph{and} it tunes the replicon, $\repl(\beta)\to0$ at a finite
$\beta_{AT}$. On the $(\beta,\eps)$ manifold both coordinates are conjugate fields, so the metric is a genuine
covariance, positive-definite by construction. The receipt confirms it at \emph{two} operating points in the
finite-$T$ RSB window:
\begin{equation}
\detg>0\ \text{throughout and diverging},\qquad
|\Ric|\sim\chi^2\sim\frac{1}{(\beta_{AT}-\beta)^2},\qquad
|\Ric|\cdot(\beta_{AT}-\beta)^2\to
\begin{cases}11.8 & (\kappa=-0.5,\ \alpha=4.2,\ \beta_{AT}=5.99)\\[2pt]
12.9 & (\kappa=-0.3,\ \alpha=3.2,\ \beta_{AT}=6.32)\end{cases}
\label{eq:percR}
\end{equation}
so the $\chi^2$ law is robust across the line, not a one-point fit. This is a diverging metric
\emph{component} ($g_{\eps\eps}=\chi\to\infty$) on a \emph{non-degenerate, positive-definite} metric---the
genuine side of Definition~\ref{def:diag} in full, the literal Riemannian twin of \eqref{eq:sk}. \textbf{Curvature-genuine for the perceptron storage/jamming archetype is verified}
$[V]$ \cite[receipt \texttt{module\_L\_perceptron\_finiteT.py}; the indefinite $(\alpha,\eps)$ first pass is
\texttt{module\_L\_perceptron\_curvature.py}]{repo}.

% =====================================================================
\section{The genuine gate and the topological trichotomy}\label{sec:gate}
The three archetypes of Secs.~\ref{sec:ridge}--\ref{sec:perceptron} are \emph{non-singular}: their Fisher
metrics are non-degenerate away from the transition, and the diagnostic classifies them cleanly. Real
overparametrized networks are the \emph{singular} case --- non-identifiable, with positive-dimensional
degeneracies on which $\det I(w)\to0$ --- so whether the genuine-side construction survives there is a
separate question, decided on paper before any sampler (the analysis is archived with the receipts
\cite{repo}). This section states the gate, refutes an unconditional closure, and reduces the genuine side to
a structural trichotomy.

\subsection{From kinematic to genuine}\label{sec:gate-criterion}
Section~\ref{sec:ridge} established that double descent is kinematic: a covariance degeneracy under which the
scalar curvature stays bounded despite a diverging volume. We ask the converse --- under what conditions is a
curvature divergence \emph{genuine}, a signature of interacting criticality rather than a coordinate or
volume artifact? For the solvable archetypes the criterion is sharp (Definition~\ref{def:diag}): genuine
criticality requires a diverging metric \emph{component} on a non-degenerate, positive-definite thermodynamic
metric. Closing the genuine side for a \emph{singular} posterior therefore means showing the singular
structure forces $\detg\to0$ at the transition.

\subsection{Refutation of unconditional closure}\label{sec:gate-race}
It does not, in general. A singular minimum has density of states $\rho(E)\sim E^{\lambda-1}$, with $\lambda$
the RLCT --- the same exponent that governs the Bayes free energy $F_n\approx\lambda\log n$. Then
$Z(\beta)=\Gamma(\lambda)\,\beta^{-\lambda}$, so $\mathrm{Var}(E)=\lambda/\beta^2$ and the energy-fluctuation
component is suppressed, $g_{\beta\beta}\sim\lambda/\beta^2\to0$ --- but only in the singular
($\beta\to\infty$) limit. The curvature is then governed by the competition
\begin{equation}
\detg=g_{\beta\beta}\,g_{\eps\eps}-g_{\beta\eps}^2\ \sim\ \frac{\lambda}{\beta^2}\,\chi-g_{\beta\eps}^2,
\label{eq:detg-race}
\end{equation}
between this singular suppression and the overlap susceptibility $\chi=g_{\eps\eps}$ that drives genuine
criticality at the de Almeida--Thouless line. At a \emph{finite}-$\beta$ transition $g_{\beta\beta}=O(1)$
while $\chi\to\infty$, so $\detg\to+\infty$ and the genuine signature survives; the suppression, a
$\beta\to\infty$ effect, never reaches it. The closure can bite only where the criticality \emph{co-locates}
with the singular limit, $\chi\sim\beta^{p}$: then $\detg\sim\lambda\,\beta^{p-2}-g_{\beta\eps}^2$, and the
threshold is $p=2$ --- $p<2$ closes the gate ($\detg\to0$, F-spurious), $p>2$ leaves it open, $p=2$ is
borderline. Here $p$ is the \emph{susceptibility exponent} in $\beta$, not the order of any $p$-spin
interaction. (This corrects a first pass that claimed unconditional closure; the finite-$\beta$ case refutes
it.)

\subsection{The topological trichotomy}\label{sec:gate-trichotomy}
Because $\chi=N\,\mathrm{Var}(q)$ diverges \emph{only} at a continuous-RSB transition, the genuine side is
decided not by a rate but by \emph{where} that transition sits relative to the singular limit --- a
structural question with exactly three branches (Table~\ref{tab:trichotomy}).

\begin{table}[h]\centering
\begin{tabular}{@{}llll@{}}
\toprule
branch & RSB structure & $\detg$ as critical & verdict\\
\midrule
1 & finite-$\beta$ RSB ($\beta_{AT}<\infty$) & $\to+\infty$ ($g_{\beta\beta}=O(1)$, $\chi\to\infty$) & genuine survives; gate OPEN\\
2 & zero-$T$ RSB ($\beta\to\infty$), $\chi\sim\beta^{p}$ & $\sim\lambda\,\beta^{p-2}-g_{\beta\eps}^2$ & $p<2$ CLOSED / $p>2$ OPEN / $p=2$ borderline\\
3 & no RSB (connected manifold) & --- ($\chi$ saturates, finite) & vacuous; closed for \emph{absence} of criticality\\
\bottomrule
\end{tabular}
\caption{The genuine-side trichotomy for a singular posterior. Branch~1 is realized by the spherical
perceptron (Sec.~\ref{sec:perceptron}) and SK at the dAT line (Sec.~\ref{sec:sk}) --- which is exactly why
their genuine results are $[V]$ and a control was even conceivable. Branch~2 has no exactly-solvable
archetype here, and the model for $\chi(\beta)$ that would fix $p$ is left open. Branch~3 is the
mode-connected case of \S\ref{sec:gate-handoff}.}\label{tab:trichotomy}
\end{table}

In Branch~1 the singular suppression is irrelevant --- it lives at $\beta\to\infty$, the transition is at
finite $\beta$ --- so the genuine signature is intact, at the cost that an equilibrium sampler is then needed
after all. Branch~2 is the only branch where the \eqref{eq:detg-race} race is decisive. In Branch~3 two
replicas spread over a connected manifold, so $\chi$ saturates to a finite, $\beta$-independent constant and
never diverges: the genuine signal is vacuous, the gate closed not by suppression but by the \emph{absence}
of a transition to detect. Whether a real network's posterior is Branch~1, 2, or 3 is the spin-glass-of-nets
question, addressed empirically next.

\subsection{The handoff (conditional)}\label{sec:gate-handoff}
Empirically and, for wide networks, provably, trained-network solution sets are connected --- by low-loss
curves \cite{garipov2018,draxler2018}, as connected sublevel sets under over-parametrization
\cite{nguyen2019}, and linearly once permutation symmetry is quotiented \cite{entezari2022,ainsworth2023}.
Connectivity is the geometric signature of absent ergodicity breaking, placing such posteriors in Branch~3,
where the overlap susceptibility does not diverge and the genuine Ruppeiner signature is vacuous. The
spin-glass picture of network landscapes \cite{choromanska2015} is an analogy under strong assumptions; the
most direct dynamical comparison finds true glassiness only in the under-parametrized regime
\cite{baityjesi2018}. We therefore \emph{argue} --- not prove --- that for over-parametrized networks the
genuine-criticality construction is vacuous; whether any real posterior instead realizes a finite-$\beta$
(Branch~1) or zero-$T$ (Branch~2) transition remains open. Where the signature is vacuous, the appropriate
invariant is the singularity order --- the RLCT \cite{watanabe2009}, estimated as the LLC \cite{lau2024}. Two
distinct facts co-locate on one
object: the connected singular set is where the Ruppeiner signature goes vacuous (no RSB) and,
simultaneously, where the Fisher information degenerates ($\det I\to0$); it is the latter that makes the LLC
the natural measurement there. Two facts, one object, not one mechanism --- which is why singular learning
theory measures the LLC in exactly this regime.

% =====================================================================
\subsection{An empirical probe: the branch placement, bounded by measurement}\label{sec:gate-empirical}
The handoff of \S\ref{sec:gate-handoff} is argued. We close with a first \emph{empirical} probe of the branch
placement on a samplable surrogate --- not a settlement, but a measured bound where the posterior is reachable.

\paragraph{Setup.} A two-layer $\tanh$ network (width $H$, scalar output) on an under-determined
teacher--student task ($n=24\ll$ parameters) --- the regime where the question is non-trivial, since a
realizable, identifiable task collapses to a unique solution and a vacuous $q\to1$. The tempered posterior
$\pi_\beta\propto\pi_0(w)\,e^{-\beta nL(w)}$ (prior $\pi_0$ as base measure, likelihood tempered, so
$\beta\to0$ is a proper melt) is sampled by Metropolis-adjusted Langevin with parallel tempering on a ladder
spaced equally in thermodynamic length $\mathcal L=\int\sqrt{C(\beta)}\,d\ln\beta$, $C=\beta^2\mathrm{Var}(E)$;
the rung count is thereby an \emph{output}, not a guess. The order parameter is the \emph{function-space}
overlap $q$ of two posterior replicas on a fixed holdout probe --- permutation-invariant by construction, so
weight-space symmetry cannot counterfeit it --- with susceptibility $\chi_F=N\,\mathrm{Var}(q)$.

\paragraph{Certification.} A width is admitted only on evidence that the sampler sees the whole posterior, not
a sub-region: (i)~two replicas \emph{warm-started from independently-trained solutions}, verified separated
($q_{\mathrm{init}}<0.9$), must converge to a common cold-$q$ distribution --- init-independence read as a
direct connectivity test; (ii)~tempering round-trip flow exceeding one per replica with the inter-init flow
spread closed; (iii)~function-space $\hat R<1.1$. Widths failing any gate are recorded \texttt{INCONCLUSIVE}
and not read.

\paragraph{Result.} Across the reachable widths the susceptibility \emph{grows}:
$\chi_F=0.54\,(H{=}32)\to0.99\,(H{=}64)$, both certified (separated replicas converged, $dq\sim10^{-3}$; flow
spread $\le0.03$). At $H{=}128$ the machinery \emph{cannot certify} --- round-trip flow stalls below one per
replica even on the thermodynamic-length ladder --- so $\chi_F\approx2.2$ there is inadmissible. This is a
\emph{bound}, not a null: the budget at which $H{=}128$ stalls is a quantitative floor on the mixing barrier,
the genuinely-glassy corner where the branch-2/branch-3 distinction would be decided. Two controls discipline
the read. An earlier, \emph{un}separated-init certification of $H{=}128$ was \emph{refuted} by the adversarial
test (its flow was init-dependent), confirming that the basin door must be shut before any growth is read.
And the energy-side specific heat stays \emph{broad} at every width (no sharpening peak), so the overlap-side
growth is \emph{not} corroborated by a thermodynamic-transition signature.

\paragraph{What it bounds.} The branch question is now bounded by measurement rather than argument: a
certified, basin-verified $\chi_F$ growth where the posterior is reachable; provably \emph{un}reachable at the
width where branch~2 and branch~3 would separate; and \emph{uncorroborated} by the specific heat. The
reachable evidence is ambiguous between sub-critical growth without a transition and a transition that cannot
be sampled --- the absent specific-heat peak leans the former, mildly consistent with the mode-connectivity
(branch~3) reading of \S\ref{sec:gate-handoff}, but as a lean, not a settlement. Status \texttt{[E-hw]}: one
task and architecture family, two certified widths; the receipts and the adversarial-certification protocol
are archived with this work \cite{repo}. A stronger glassy-regime sampler for $H\ge128$ and a denser width
sweep are the deliberate next passes, neither required for the bound stated here.

% =====================================================================
\section{Discussion}\label{sec:discussion}
The four-tier picture (Table~\ref{tab:tax}) is uniform across venues. A divergence earns ``genuine'' only as
a positive-definite component blow-up; the metric-volume explosion of double descent, the
coordinate-degeneracy of Curie--Weiss, and the blindness of a manifold without the conjugate field are its
three foils.

\begin{table}[h]\centering
\begin{tabular}{@{}llll@{}}
\toprule
system (venue) & metric $\detg$ & scalar $\Ric$ & verdict\\
\midrule
ideal Bose gas (BEC) & non-degenerate, diverging comp. & $|\Ric|\to\infty$ & genuine (classical \cite{janyszek1990})\\
ridge / double descent (learning) & $\to\infty$ (rank-1) & $\to0^-$ (noiseless) / bdd $<0$ (noisy) & kinematic\\
Curie--Weiss, dependent coords & $\to0$ / indefinite & spurious $|\Ric|\to\infty$ & degenerate\\
bare SK $(\beta,h)$ & finite & bounded & blind\\
$\eps$-augmented SK at dAT (spin glass) & $>0$, diverging comp. & $|\Ric|\to\infty$ & genuine $[V]$\\
spherical perceptron, $\kappa\ge0$ (storage) & regular & bounded & kinematic\\
spherical perceptron, $\kappa<0$ (jamming) & $>0$, diverging comp. & $|\Ric|\sim\chi^2$ & genuine $[V]$\\
\bottomrule
\end{tabular}
\caption{The diagnostic across venues. The perceptron uniquely supplies both a kinematic and a genuine
entry.}\label{tab:tax}
\end{table}

The ideal Bose gas at condensation (Table~\ref{tab:tax}, first row) is the canonical classical exemplar of
the genuine signature \cite{janyszek1990}, included as the literature anchor; the three disordered-system
archetypes below it are this paper's contribution.

The organising claim is that none of these systems is special: the dichotomy is a statement about disordered
systems with a continuous replicon, and learning is merely the venue in which the question is most often
asked. None of the underlying phase structure is new---it is the classical content of
\cite{gardner1988,gardnerderrida1988,krauthmezard1989,sk1975,at1978}. What is verified ($[V]$) is the
per-system curvature behaviour of each archetype; their \emph{uniformity under a single criterion}---that
these facts collectively constitute a general diagnostic separating genuine from kinematic---is the
extrapolation we propose, supported by but not reducible to the three exactly-solvable cases. The discipline
of Definition~\ref{def:diag} (positive-definite signature, not merely $|\Ric|\to\infty$) is load-bearing
rather than pedantic: the $(\alpha,\eps)$ versus $(\beta,\eps)$ contrast of Sec.~\ref{sec:perc-R} is exactly
a case where the wrong coordinate produces a divergence that \emph{looks} genuine on an indefinite metric and
is not.

\paragraph{Related work.} The closest published prior art is the curvature analysis of $L_2$-regularized
deep networks by Ersoy and Wiesner \cite{ersoy2025}, who explain the regularization-driven \emph{onset of
learning}---a first-order transition into the under-parametrized phase, with further transition points and
hysteresis as data complexity grows, and a reinterpretation of ``grokking''---through the scalar (Ricci)
curvature of the \emph{error landscape}, extending the depth-dependent transition orders (second order for a
single hidden layer, first order for deeper nets) of Ziyin and Ueda \cite{ziyinueda2023}. The overlap is
real, but the geometric objects are distinct. (i)~\emph{The metric.} Theirs is the embedding geometry of a
fixed loss surface (its first and second fundamental forms; Gauss--Kronecker and Ricci scalars); ours is the
Ruppeiner thermodynamic metric \eqref{eq:metric}, the Hessian of a disorder-averaged free energy---a
\emph{covariance of fluctuations}. A deterministic loss surface carries no fluctuation content, so it cannot
make the ``$|\Ric|\sim$ correlation volume'' reading that is the center of our diagnostic. (ii)~\emph{The
control field.} Theirs is the $L_2$ penalty strength (which they also denote $\beta$); ours is a genuine
inverse temperature, with the conjugate coordinate $\eps$ the order parameter's, not a regularizer's.
(iii)~\emph{Disorder.} Their construction is a single deterministic realization, with no quenched average and
no replicon; our kinematic-versus-genuine distinction turns on the stability of the replica-symmetric
saddle---the replicon / de Almeida--Thouless line---which their setting does not contain. (iv)~\emph{The
taxonomy.} Their first-/second-order classification (after Ziyin--Ueda) concerns the \emph{continuity} of the
collapse; ours (Definition~\ref{def:diag}) concerns whether a curvature divergence is genuine interacting
criticality or a \emph{kinematic} volume divergence---an orthogonal axis, since a transition that is
second-order in their sense may be either kinematic or genuine in ours, a distinction their single-realization
metric cannot draw. (v)~\emph{Degeneracy.} Where they observe that the Ricci scalar is numerically more
stable than the Gauss--Kronecker curvature, our claims rest on \emph{establishing} a non-degenerate,
positive-definite metric on the relevant submanifold, without which $|\Ric|\to\infty$ is disqualified as a
coordinate artifact. (vi)~\emph{Scope.} Their results are numerical, on small networks; ours are closed-form
on exactly-solvable mean-field archetypes (SK, the spherical perceptron) with exact capacities and AT lines.
The two are complementary: theirs measures the shape of a deterministic loss surface, ours the fluctuation
geometry of a disordered free energy, and the genuine-versus-kinematic distinction is not available on their
object.

\paragraph{Limitations.} The diagnostic presumes one can identify and adjoin the order parameter's conjugate
field $\eps$; absent that, the manifold is blind by construction, so the method \emph{classifies} known
transitions rather than discovering order parameters. Our validation is on mean-field, replica-symmetric (or
RS-approached) exactly-solvable archetypes on two-parameter manifolds; extension to finite-dimensional
systems, to full-RSB structure, and to higher-dimensional manifolds is open.

% =====================================================================
\appendix
\section{The curvature engine and receipts}\label{app:engine}
All curvatures use $\Ric=-N/(2(\detg)^2)$ with $N=g_{xx}(g_{xxy}g_{yyy}-g_{xyy}^2)
-g_{xy}(g_{xxx}g_{yyy}-g_{xyy}g_{xxy})+g_{yy}(g_{xxx}g_{xyy}-g_{xxy}^2)$, pinned to $\Ric=-1$ on the normal
family. Five computational receipts reproduce every numerical claim:
\texttt{module\_L\_ridge\_curvature.py} (Sec.~\ref{sec:ridge}),
\texttt{module\_L\_SK\_converse.py} (Sec.~\ref{sec:sk}),
\texttt{module\_L\_perceptron\_replicon.py} (the susceptibility, Sec.~\ref{sec:perc-chi}),
\texttt{module\_L\_perceptron\_finiteT.py} (the positive-definite $(\beta,\eps)$ curvature, Sec.~\ref{sec:perc-R}),
and the indefinite-metric first pass \texttt{module\_L\_perceptron\_curvature.py} (the cautionary
$(\alpha,\eps)$ intermediate). The anchors $\alpha_c(0)=2$ and $s(0)=\log2$ are exact to machine precision,
the $(\beta,\eps)$ replicon is anchored to Gardner ($\beta\!\to\!\infty$) and the engine to $\Ric=-1$.

\paragraph{Reproducibility.} All receipts are CPU-only Python (\texttt{numpy}/\texttt{scipy}; \texttt{mpmath}
at 40 decimal places where cancellation requires it, e.g.\ the rank-1 ridge metric). Gaussian integrals use
$200$-node Gauss--Hermite quadrature; saddles are solved by \texttt{brentq}/\texttt{fsolve} to tolerance
$10^{-13}$; finite differences for $\Ric$ are used only where Remark~\ref{rem:fd} certifies them, with the
load-bearing components obtained by closed form or implicit-differentiation linear response. The receipts and
the gate document \texttt{T1\_moduleL\_perceptron\_gate} are archived with this work \cite{repo}.

% =====================================================================
\begin{thebibliography}{99}
\bibitem{ruppeiner1995} G.~Ruppeiner, \emph{Riemannian geometry in thermodynamic fluctuation theory},
  Rev.\ Mod.\ Phys.\ \textbf{67}, 605 (1995); erratum \emph{ibid.} \textbf{68}, 313 (1996).
\bibitem{janyszek1990} H.~Janyszek and R.~Mruga\l a, \emph{Riemannian geometry and the stability of ideal
  quantum gases}, J.\ Phys.\ A \textbf{23}, 467 (1990).
\bibitem{belkin2019} M.~Belkin, D.~Hsu, S.~Ma, S.~Mandal, \emph{Reconciling modern machine-learning practice
  and the classical bias--variance trade-off}, PNAS \textbf{116}, 15849 (2019).
\bibitem{sk1975} D.~Sherrington and S.~Kirkpatrick, \emph{Solvable model of a spin-glass}, Phys.\ Rev.\
  Lett.\ \textbf{35}, 1792 (1975).
\bibitem{at1978} J.~R.~L.~de Almeida and D.~J.~Thouless, \emph{Stability of the Sherrington--Kirkpatrick
  solution of a spin glass model}, J.\ Phys.\ A \textbf{11}, 983 (1978).
\bibitem{gardner1988} E.~Gardner, \emph{The space of interactions in neural network models}, J.\ Phys.\ A
  \textbf{21}, 257 (1988).
\bibitem{gardnerderrida1988} E.~Gardner and B.~Derrida, \emph{Optimal storage properties of neural network
  models}, J.\ Phys.\ A \textbf{21}, 271 (1988).
\bibitem{krauthmezard1989} W.~Krauth and M.~M\'ezard, \emph{Storage capacity of memory networks with binary
  couplings}, J.\ Phys.\ France \textbf{50}, 3057 (1989).
\bibitem{gyorgyi1990} G.~Gy\"orgyi, \emph{First-order transition to perfect generalization in a neural
  network with binary synapses}, Phys.\ Rev.\ A \textbf{41}, 7097 (1990).
\bibitem{ziyinueda2023} L.~Ziyin and M.~Ueda, \emph{Zeroth, first, and second-order phase transitions in
  deep neural networks}, Phys.\ Rev.\ Research \textbf{5}, 043243 (2023).
\bibitem{ersoy2025} I.~T.~Ersoy and K.~Wiesner, \emph{Phase Transitions between Accuracy Regimes in $L_2$
  regularized Deep Neural Networks}, arXiv:2505.06597 (2025).
\bibitem{garipov2018} T.~Garipov, P.~Izmailov, D.~Podoprikhin, D.~Vetrov, A.~G.~Wilson, \emph{Loss surfaces,
  mode connectivity, and fast ensembling of DNNs}, Adv.\ Neural Inf.\ Process.\ Syst.\ (NeurIPS) \textbf{31},
  8803 (2018); arXiv:1802.10026.
\bibitem{draxler2018} F.~Draxler, K.~Veschgini, M.~Salmhofer, F.~A.~Hamprecht, \emph{Essentially no barriers
  in neural network energy landscape}, Proc.\ 35th ICML, PMLR \textbf{80}, 1308 (2018); arXiv:1803.00885.
\bibitem{nguyen2019} Q.~Nguyen, \emph{On connected sublevel sets in deep learning}, Proc.\ 36th ICML, PMLR
  \textbf{97}, 4790 (2019); arXiv:1901.07417.
\bibitem{entezari2022} R.~Entezari, H.~Sedghi, O.~Saukh, B.~Neyshabur, \emph{The role of permutation
  invariance in linear mode connectivity of neural networks}, Int.\ Conf.\ Learning Representations (ICLR)
  (2022); arXiv:2110.06296.
\bibitem{ainsworth2023} S.~K.~Ainsworth, J.~Hayase, S.~Srinivasa, \emph{Git Re-Basin: merging models modulo
  permutation symmetries}, Int.\ Conf.\ Learning Representations (ICLR) (2023); arXiv:2209.04836.
\bibitem{choromanska2015} A.~Choromanska, M.~Henaff, M.~Mathieu, G.~Ben~Arous, Y.~LeCun, \emph{The loss
  surfaces of multilayer networks}, Proc.\ 18th AISTATS, PMLR \textbf{38}, 192 (2015); arXiv:1412.0233.
\bibitem{baityjesi2018} M.~Baity-Jesi, L.~Sagun, M.~Geiger, S.~Spigler, G.~Ben~Arous, C.~Cammarota, Y.~LeCun,
  M.~Wyart, G.~Biroli, \emph{Comparing dynamics: deep neural networks versus glassy systems}, Proc.\ 35th
  ICML, PMLR \textbf{80}, 324 (2018); arXiv:1803.06969.
\bibitem{watanabe2009} S.~Watanabe, \emph{Algebraic Geometry and Statistical Learning Theory}, Cambridge
  Univ.\ Press (2009).
\bibitem{lau2024} E.~Lau, Z.~Furman, G.~Wang, D.~Murfet, S.~Wei, \emph{The local learning coefficient: a
  singularity-aware complexity measure}, arXiv:2308.12108 (2024).
\bibitem{repo} IG-PRIMON-T1 computational receipts (this work);
  \texttt{module\_L\_*.py}, gate document \texttt{T1\_moduleL\_perceptron\_gate\_v0\_1.md}.
\end{thebibliography}

\end{document}
