\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,geometry,booktabs,hyperref}
\geometry{margin=1.1in}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newcommand{\eps}{\varepsilon}
\title{The Fisher metric of the primon gas at the Hagedorn point,\\
and a two-route instrument for the Keiper--Li coefficients\\
\large (IG-PRIMON-T1 note, draft v0.1 --- Phase 1 of the registration of 2026-06-12)}
\author{B.~Jones\thanks{Computations performed on consumer hardware; all receipts, pre-registrations,
and gate records in the accompanying repository. Draft for read-through; no novelty wording final
until all literature gates are recorded.}}
\date{2026-07-08}
\begin{document}
\maketitle

\begin{abstract}
We study the Fisher information $I(\beta)=\operatorname{Var}(\log n)=(\log\zeta)''(\beta)$ of the
Gibbs family $p_n(\beta)\propto n^{-\beta}$ (the ``primon gas'') near its Hagedorn point $\beta=1$.
We record the expansion $I(1+\eps)=\eps^{-2}+\sum_k c_k\eps^k$ with $c_0,c_1$ in closed Stieltjes
form and $c_k$ verified to $50$ digits ($k\le 6$); we prove the radius of convergence is exactly $3$
and that the $c_k$ are shifted power sums over the full Riemann zero set --- a known corollary of the
Hadamard product (Lehmer; Keiper; Bombieri--Lagarias; Coffey; Voros), which we use, not claim. The
information length to the Hagedorn point diverges logarithmically with renormalization constant
$C=-0.03435615417912198608\ldots$ ($80$ digits verified). We then run the cumulant chain upward to
the Keiper--Li coefficients: a zero-free route (contour Taylor coefficients of $\log[(s-1)\zeta(s)]$,
exact series logarithm, binomial transform) against a zero-sum route (Odlyzko ordinates plus a
semi-analytic Riemann--von Mangoldt tail with explicit budget), agreeing to $\sim\!2\times10^{-8}$
relative at $n=400$. The zero-free route's precision cost is measured as an empirical law
($\text{digits lost}\approx 0.295\,n$), and the instrument's sensitivity to hypothetical off-line
zeros is charted by planted-violation experiments. No claim of RH evidence is made or implied;
$\lambda_n>0$ in our range is occupied territory. An appendix records negative results and the
program's pre-registered falsifiers.
\end{abstract}

\section{Setup and standing discipline}
For $\beta>1$ let $p_n(\beta)=n^{-\beta}/\zeta(\beta)$. Its Fisher information is
$I(\beta)=(\log\zeta)''(\beta)$, divergent as $\beta\downarrow 1$ (the Hagedorn point). Prior art
absorbed per the registration: primon-gas thermodynamics and the Hagedorn divergence are classical
(Julia; Spector; Bakas--Bowick); Bost--Connes symmetry breaking at $\beta=1$; the Knauf chain;
Riemann zeros as Fisher/Lee--Yang zeros at physics rigor (Julia; Dutta--Ghoshal). The bare statement
$I\sim(\beta-1)^{-2}$ is the Hagedorn divergence relabeled and carries no standalone value. Every
numerical claim below carries a stated precision, an independent second route, and a falsifier.

\section{The expansion at the Hagedorn point}
Write $Z(s)=(s-1)\zeta(s)$, entire of order 1, $Z(1)=1$, and $S(\eps)=(\log Z)''(1+\eps)$, so that
$I(1+\eps)=\eps^{-2}+S(\eps)$, $S(\eps)=\sum_{k\ge0}c_k\eps^k$.

\textbf{Constants (locked 2026-06-12, dps 50, two independent routes, falsifier $10^{-40}$).}
\begin{align*}
c_0&=-(\gamma_0^2+2\gamma_1)=-0.18754\,62328\,40365\,22459\,72033\,84605\,44158\,83839\ldots\\
c_1&=6\left(\tfrac{\gamma_2}{2}+\gamma_0\gamma_1+\tfrac{\gamma_0^3}{3}\right)
=+0.10337\,72640\,66385\,78760\,40164\,46167\,20832\,68908\ldots
\end{align*}
with $c_2,\dots,c_6$ tabulated in the repository at the same precision.

\begin{lemma}[radius and zero-sum representation; proof in the repository lemma file]
(i) $(\log Z)''(s)=-\sum_\rho (s-\rho)^{-2}$ over all zeros of $\zeta$ (trivial and nontrivial,
conjugate-paired), absolutely and locally uniformly. (ii) $c_k=-(k+1)\sum_\rho(\rho-1)^{-(k+2)}$.
(iii) The radius of convergence of $\sum c_k\eps^k$ is exactly $3$, with
$c_k\sim-(k+1)(-3)^{-(k+2)}$ and the entire nontrivial-zero contribution $O(k\cdot 14.13^{-k})$.
\end{lemma}

\textbf{Attribution (per the executed literature gate).} Part (ii) is known-as-stated in equivalent
forms: Lehmer (1988) at $s=0$; Keiper (1992) for $\log\xi$ at $s=1$; Bombieri--Lagarias (1999),
\S4 (Stieltjes cumulants); Coffey (2005), Thm.~3 ($\eta$-coordinates); Voros (2005), eqs.~(49)--(56)
(arbitrary expansion point, general $L$). We record the $c_k$-normalization for convenience only.
Direct numerical corroboration: zero-sums over the first 120 nontrivial zeros reproduce $c_2,c_4,c_6$
with residuals at $0.98$--$0.99$ of the analytic tail budget; and the derived
$\sigma_3=c_1/2-\lambda(3)+1=-1.11158231452105922762668\times10^{-4}$ matches Lehmer's Table~5 to
all 25 computed digits --- a cumulant-route/zero-route agreement across a 38-year-old table.

\section{Information length to the Hagedorn point}
$L(\beta)=\int_\beta^2\sqrt{I(t)}\,dt=-\log(\beta-1)+C+o(1)$: the Hagedorn point lies at infinite
Fisher--Rao distance, logarithmically, with
\[
C=-0.03435615417912198608311088145844761509924587622554538185676859393077026298076215\!\ldots
\]
(80 digits; tanh-sinh and Gauss--Legendre agreement at dps 100; cancellation-free integrand assembly
recorded in the gate file). Falsifier: any independent evaluation outside $10^{-81}$.

\section{From cumulants to Keiper--Li: a two-route instrument}
Li's criterion: RH $\iff \lambda_n\ge0\ \forall n$, $\lambda_n=\sum_\rho[1-(1-1/\rho)^n]$.
We compute $\lambda_n$ two independent ways and require budgeted agreement.

\textbf{Route A (zeros).} Sum over the first $10^5$ Odlyzko ordinates plus a semi-analytic tail
(integral of the pair term against the Riemann--von Mangoldt density), with budget = $S(T)$
fluctuation at the cut + ordinate-accuracy propagation + arithmetic floor.

\textbf{Route B (zero-free).} Trapezoid-DFT Taylor coefficients of the \emph{entire} $Z(1+\eps)$ on
$|\eps|=3$ (no branch hazard), exact power-series logarithm by recurrence, $\sigma_j$ via the closed
trivial-zero part, binomial transform to $\lambda_n$. Anchors: $\lambda_1=1+\gamma/2-\tfrac12\log4\pi$
(exact to working precision); $\sigma_2,\sigma_3$ vs.\ Lehmer; cross-run $\lambda_{40}$ through
independent contour machinery.

\textbf{Receipts.} At dps 130, routes agree at $0.2$--$0.3$ of budget for $n=50..400$
($\lambda_{400}=748.31558\ldots$, $|A-B|\approx1.7\times10^{-5}$, $\sim2\times10^{-8}$ relative).
An earlier version of the Route-A budget missed the $\binom{n}{2}$ tail term; the gate's failure
mode (an exact $n^2$ excess) located it --- recorded, not hidden.

\textbf{The wall, measured --- and identified as Johansson's law.} The zero-free route loses
precision linearly: $\text{digits lost}\approx 0.2947\,n$ ($R^2=0.9999$ on the (90,130)-dps pair;
confirmed out-of-sample at $0.2990$, $R^2=1.0000$, on the (310,360) pair; a contaminated first fit
and a wrong a-priori prediction are both recorded). Per the executed literature gate this is a blind
replication of a published law: Johansson prices the route at ``about $n$ bits'' ($1$ bit
$\approx0.301$ digits per step), and Ma\'slanka's Fig.~6 shows $\approx0.2$ digits/step; Voros
aggregates both. We claim only the corroboration and the explicit costing form
$n_{\mathrm{wall}}(D)\approx(D+3)/0.295$ used to price our own runs (receipts to $n=1000$ at
dps~360/310, agreement at $0.20$ of budget uniformly, $\lambda_{1000}=2326.05316\ldots$ to
$\sim5\times10^{-8}$ relative).

\textbf{Violation sensitivity (Voros's thresholds, empirically confirmed).} Planting synthetic
off-line zero quadruples: at height $\gamma^*=2$, strength $\beta^*=0.9$ turns $\lambda_n$ negative
by $n=26$; $\beta^*=0.6$ by $n=231$; sensitivity falls superquadratically with height
($\gamma^*=14.13$ invisible at $n\le400$). This is the empirical face of Voros's published RH-false
asymptotics and detection thresholds ($n\gtrsim T^2/|t|$ onset; demonstrated by him on the genuine
off-line zeros of Davenport--Heilbronn); our chart confirms, and claims, nothing beyond the
confirmation. Moreover, by Oesterl\'e's bound (RH verified to height $T_0$ $\Rightarrow$
$\lambda_n\ge0$ for $n\le T_0^2$), positivity in every range we touch is \emph{guaranteed} by
existing zero verification --- our $\lambda_n>0$ observations are consistency checks of the
instrument, not information about RH.

\section{No-go clarifications (inoculation)}
(i) The complex-$\beta$ zeros of the continued partition function are the nontrivial zeros of
$\zeta$; they are isolated, off the real axis, with no real-axis accumulation: no Lee--Yang/Fisher
transition is associated with them (Dutta--Ghoshal at physics rigor; cited, not rediscovered).
(ii) Matching a random-matrix universality class identifies nothing arithmetic: GUE statistics of
zero spacings are the anonymous fixed point; the primes live in the \emph{departures} (explicit-formula
residue), as our point-cloud demonstration (spectral lines at $x=\log p^k$, 12/12) illustrates.
(iii) One-parameter ``curvature'' claims are vacuous (1-D manifolds are flat); nothing here uses them.

\section{Honest scope}
No statement in this note constitutes evidence for or against RH; by Oesterl\'e's bound our
positivity observations were guaranteed in advance. All $\lambda_n$ values are within previously
computed ranges (Keiper to $7000$; Johansson, rigorously, to $10^5$). The final literature gate
(executed 2026-07-08, recorded in the repository) demoted all three candidate-novel items: the
two-route cross-verification is in essence Keiper (1992, including his unused hybrid formula (48))
plus the Keiper--Johansson cross-check; the wall law is Johansson's one-bit-per-step statement
(anticipated by Ma\'slanka's Fig.~6); the sensitivity chart empirically confirms Voros's published
thresholds. What remains ours is presentational only --- the per-row matched-budget receipt format
--- and the corroborations themselves, which we regard as validation of the instrument rather than
contribution to the mathematics. Negative results and all failed instrument versions are recorded in
the repository gate files; the pre-registrations bind wording.

\begin{thebibliography}{19}
\bibitem{Julia} B.~Julia, Statistical theory of numbers, in \emph{Number Theory and Physics}, Springer (1990).
\bibitem{Spector} D.~Spector, Supersymmetry and the M\"obius inversion function, Commun.\ Math.\ Phys.\ \textbf{127} (1990).
\bibitem{BakasBowick} I.~Bakas, M.~Bowick, Curiosities of arithmetic gases, J.~Math.\ Phys.\ \textbf{32} (1991).
\bibitem{BostConnes} J.-B.~Bost, A.~Connes, Hecke algebras, type III factors and phase transitions\ldots, Selecta Math.\ \textbf{1} (1995).
\bibitem{Knauf} A.~Knauf, On a ferromagnetic spin chain, Commun.\ Math.\ Phys.\ \textbf{153} (1993).
\bibitem{DuttaGhoshal} S.~Dutta, D.~Ghoshal, Phase transitions of the Riemann gas\ldots, arXiv:2102.13445.
\bibitem{Ruppeiner} G.~Ruppeiner, Riemannian geometry in thermodynamic fluctuation theory, Rev.\ Mod.\ Phys.\ \textbf{67} (1995).
\bibitem{Lehmer} D.~H.~Lehmer, The sum of like powers of the zeros of the Riemann zeta function, Math.\ Comp.\ \textbf{50} (1988).
\bibitem{Keiper} J.~B.~Keiper, Power series expansions of Riemann's $\xi$ function, Math.\ Comp.\ \textbf{58} (1992).
\bibitem{Li} X.-J.~Li, The positivity of a sequence of numbers and the Riemann hypothesis, J.~Number Theory \textbf{65} (1997).
\bibitem{BL} E.~Bombieri, J.~C.~Lagarias, Complements to Li's criterion\ldots, J.~Number Theory \textbf{77} (1999).
\bibitem{Coffey} M.~W.~Coffey, Toward verification of the Riemann hypothesis\ldots, Math.\ Phys.\ Anal.\ Geom.\ \textbf{8} (2005).
\bibitem{Voros} A.~Voros, Zeta functions over zeros of general zeta and $L$-functions, in Dev.\ Math.\ \textbf{14}, Springer (2005); and \emph{Zeta Functions over Zeros of Zeta Functions}, Springer (2010).
\bibitem{Maslanka} K.~Ma\'slanka, Li's criterion for the Riemann hypothesis --- numerical approach, Opuscula Math.\ (2004).
\bibitem{Johansson} F.~Johansson, Rigorous high-precision computation\ldots (Arb); Keiper--Li computations to $n=10^5$.
\bibitem{Odlyzko} A.~M.~Odlyzko, Tables of zeros of the Riemann zeta function, \url{www.dtc.umn.edu/~odlyzko}.
\end{thebibliography}
\end{document}
